### 2021-06-02 paper/paper.tex
passed a rule function to the step function of the state class

\documentclass[11pt]{article}
    \title{\textbf{Causal graph quantum dynamic simulation}}
    \author{auteurs...}
    \date{}

    \setlength{\parindent}{0cm}

    \addtolength{\topmargin}{-3cm}
    \addtolength{\textheight}{3cm}
    \usepackage[margin=0.5in]{geometry}

    \usepackage{xcolor}
	\usepackage{listings}

	\definecolor{mGreen}{rgb}{0,0.6,0}
	\definecolor{mGray}{rgb}{0.5,0.5,0.5}
	\definecolor{mPurple}{rgb}{0.58,0,0.82}
	\definecolor{backgroundColour}{rgb}{0.95,0.95,0.92}

	\lstdefinestyle{CStyle}{
		backgroundcolor=\color{backgroundColour},   
		commentstyle=\color{mGreen},
		keywordstyle=\color{magenta},
		numberstyle=\tiny\color{mGray},
		stringstyle=\color{mPurple},
		basicstyle=\footnotesize,
		breakatwhitespace=false,         
		breaklines=true,                 
		captionpos=b,                    
		keepspaces=true,                 
		numbers=left,                    
		numbersep=5pt,                  
		showspaces=false,                
		showstringspaces=false,
		showtabs=false,                  
		tabsize=2,
		language=C
	}
\begin{document}

\maketitle
\thispagestyle{empty}

\section{Introduction}

\section{Results}

\section{Tests}

We did some tests to validate the simulations, including (but not limitted to) the classical and quantum injectivity tests.

\subsection{classical injectivity test}

To check if our graph dynamics is reversible, we can apply on a huge number of random graphs a single step, and a reversed step and check that we get back the same graph with started with.\\

This was tested on arround 10k graphs, for chich the particules position were radomized, but the graph name was kept, making it more complex as each step on random particules positions caused splits and merges.

\subsection{quantum injectivity test}

To check that our graph dynamic is unitary, we can check that computing an arbitrary amount of steps on an arbitrary initial state (incrissing drasticly the number of graphs), and then apply the same number of reversed step, and we check that we get back the state we started with.\\

Even including floating points errors, we can esaily do this test with around 5 steps, obtain around 7M graphs after the 5 forward iteration, and comme back to the same step within $10^{-6}$ error in the magnitude.

\section{Implementation}

The code shown below is a simplification of the actual code  \footnote{https://github.com/jolatechno/Quantum-graph-simulation.git} (classes are shown without member function or decorators, and code inside loops is represented by functions).\\

The folllowing classes are used in the implementation:

\begin{description}
\addtolength{\itemindent}{0.80cm}
\itemsep0em 
\item[node names] corresponding to the name \footnote{arithmetique de nom} of a single node.
\item[graph names] i.e. ordered set of node.
\item[graphs with particules] graph name and positions of particules.
\item[superposition of graphs]
\end{description}

\subsection{node names}

Nodes can either be "left" or "right" containers (pointing to another node), "elemnts" (a simple integer name) or pairs (a merge of two other nodes, pointing to two other nodes).\\

We mark each node with a "has\_most\_left\_zero\_" (which is inerited through merges)  variable to keep track of the node that contains the first nodes.

\subsubsection{class}

\begin{lstlisting}[style=CStyle]
struct node {
	int left_idx__or_element_;
	int right_idx__or_type_;
	size_t hash_ = 0;
	bool has_most_left_zero_ = false;
	//...
\end{lstlisting}

\subsubsection{constructors}

The following constructor helps construct an "element-type" node (which are just a single integer, and are used to initialize graphs):

\begin{lstlisting}[style=CStyle]
	//...
	node(int n) :
		left_idx__or_element_(n), right_idx__or_type_(element_idx) {

		hash_ = n;
		boost::hash_combine(hash_, element_idx);

		has_most_left_zero_ = n == 0;
	}
	//...
\end{lstlisting}

We also need a constructor for "left/eright-type" node (noted "u.l" or "u.r"):

\begin{lstlisting}[style=CStyle]
	//...
	node(int const idx, node_t const &other, int const type) :
		left_idx__or_element_(idx), right_idx__or_type_(type) {

		boost::hash_combine(hash_, other.hash_);
		boost::hash_combine(hash_, type);

		// check for most left zero
		if (is_left())
			has_most_left_zero_ = other.has_most_left_zero_;
	}
	//...
\end{lstlisting}

And finally, we need a constructor to construct "pair-type" nodes (noted "u$\wedge$v"):

\begin{lstlisting}[style=CStyle]
	//...
	node(int const left_idx, node_t const &left,
		int const right_idx, node_t const &right) :
		left_idx__or_element_(left_idx),
		right_idx__or_type_(right_idx){

		boost::hash_combine(hash_, left.hash_);
		boost::hash_combine(hash_, right.hash_);
		has_most_left_zero_ = left.has_most_left_zero_ || right.has_most_left_zero_;
	}
	//...
}
\end{lstlisting}

\subsubsection{hash}

the hash of a node is stored in the "hash\_" varibale, and is simply calculated by the constructor (since it never changes).\\

To compute the hash of a node, we simply combine the hash of all children nodes or of the integer representing the element, and/or the type of the node.

\subsection{graph names}

To store the list of nodes, we use a main nodes list, which we keep in a lexicographical order (the first node should always be the node with the "most left zero").\\

Other nodes which are pointed to are stored into a buffer vector, for which we use grabage collection.

\subsubsection{class}

\begin{lstlisting}[style=CStyle]
class graph_name {
private:
	// node list 
	std::vector<node_t> nodes_;
	std::vector<node_t> node_buff_;
	std::vector<int> trash_collection_;

	// hash
	size_t mutable hash_ = 0;
	bool mutable hashed_ = false;

	//...
\end{lstlisting}

\subsubsection{usefull functions}

Since we use trash collection to manage memory, we need to implement a collection to insert a node into the "buffer vector":

\begin{lstlisting}[style=CStyle]
	//...
	int push_to_buffer(node_t &n) {
		if (trash_collection_.empty()) {
			node_buff_.push_back(n);
			return node_buff_.size() - 1;
		}

		int buff_idx = trash_collection_.back();
		trash_collection_.pop_back();
		node_buff_[buff_idx] = n;
		return buff_idx;
	}
	//...
\end{lstlisting}

\subsubsection{split}

\begin{lstlisting}[style=CStyle]
	//...
	bool inline split(unsigned int idx) {
		hashed_ = false;

		node_t node = nodes_[idx];
		if (node.is_pair()) {
			// read indexes
			int const left_idx = node.left_idx();
			int const right_idx = node.right_idx();

			// add left node
			nodes_[idx] = node_buff_[left_idx];
			trash_collection_.push_back(left_idx);

			// add right node
			nodes_.insert(nodes_.begin() + idx + 1, node_buff_[right_idx]);
			trash_collection_.push_back(right_idx);

			// check if we should rotate
			if (idx == 0 && nodes_[1].has_most_left_zero()) {
				std::rotate(nodes_.begin(), nodes_.begin() + 1, nodes_.end());
				return true; 
			}

			return false;
		}

		// add current node to buffer
		int buff_idx = push_to_buffer(node);

		// add left node
		nodes_[idx] = node_t(buff_idx, node, point_l_idx);

		// add right node
		nodes_.emplace(nodes_.begin() + idx + 1, buff_idx, node, point_r_idx);

		return false;
	}
	//...
\end{lstlisting}

The boolean return argument of this function is used to check if the merge requires a rotation of nodes.

\subsubsection{merges}

\begin{lstlisting}[style=CStyle]
	//...
	void inline merge(unsigned int idx) {
		hashed_ = false;

		// destination idx
		unsigned int right_idx = (idx + 1) % size();
		auto left = nodes_[idx];
		auto right = nodes_[right_idx];

		if (left.is_left() && right.is_right()) {
			int const left_left = left.left_idx();
			int const right_left = right.left_idx();

			if (node_buff_[left_left].hash() == node_buff_[right_left].hash()) {
				// trash collect
				trash_collection_.push_back(left_left);
				if (left_left != right_left)
					trash_collection_.push_back(right_left);

				//add node
				nodes_[right_idx] = node_buff_[left_left];
				nodes_.erase(nodes_.begin() + idx);
				return;
			}
		}

		//add node
		int const left_idx = push_to_buffer(left);
		int const right_idx_ = push_to_buffer(right);

		nodes_[right_idx] = node_t(left_idx, left, right_idx_, right);
		nodes_.erase(nodes_.begin() + idx);
	}
	//...
\end{lstlisting}

\subsubsection{hash()}

To hash a graph name, we simply combine the hash of all the nodes.

\begin{lstlisting}[style=CStyle]
	//...
	size_t inline hash() const {
		if (hashed_)
			return hash_;

		hash_ = 0;

		/* combine all nodes hash */
		for (auto &n : nodes_)
			boost::hash_combine(hash_, n.hash());

		hashed_ = true;
		return hash_;
	}
	//...
}
\end{lstlisting}

\subsection{graphs with particules}

A graph is represented by a graph name, and an ordered list of the node's index of particules going both ways.

\subsubsection{class}

\begin{lstlisting}[style=CStyle]
class graph {
	// variables 
	graph_name_t* name_;
	std::vector<unsigned int> left_;
	std::vector<unsigned int> right_;

	// hash
	size_t mutable hash_ = 0;
	bool mutable hashed_ = false;

	//...
\end{lstlisting}

\subsubsection{step and reversed\_step()}

The fonction to step all particules uses the fonctions "rotate\_once\_left" and "rotate\_once\_right" which rotate particules if they overflow:

\begin{lstlisting}[style=CStyle]
	//...
	void inline step() {
  		hashed_ = false;

  		rotate_once_right(right_);
  		rotate_once_left(left_);
  	}
  	//...
\end{lstlisting}

To implement the reverse stpe we simply exchange the roles of "left" and "right" in the "step" function:

\begin{lstlisting}[style=CStyle]
	//...
  	void inline reversed_step() {
  		hashed_ = false;

  		rotate_once_right(left_);
  		rotate_once_left(right_);
  	}
  	//...
\end{lstlisting}

\subsubsection{split\_merge()}

\begin{lstlisting}[style=CStyle]
	//...
	void split_merge(std::vector<split_merge_t>& split_merge) {
		if (split_merge.empty())
			return;

		hashed_ = false;

		// check for last merge
		if (split_merge.back() == split_merge_t(size() - 1, merge_t)) {
			name_->merge(size() - 1);
			overflow_right(left_);	
			split_merge.pop_back();
		}

		// check for first element split 
		bool first_split = false;

		// and calculating max displacement 
		int total_displacement = 0;
		for (auto & [pos, type] : split_merge | std::ranges::views::reverse)
			if (type == split_t) {
				++total_displacement;
			} else
				--total_displacement;

		// reserve space for nodes
		name_->reserve(total_displacement);

		// split and merge names
		for (auto & [pos, type] : split_merge | std::ranges::views::reverse)
			if (type == split_t) {
				first_split |= name_->split(pos);
			} else
				name_->merge(pos);

		// move particules
		auto const split_merge_begin = split_merge.rend();
		int displacement = total_displacement;
		auto split_merge_it = split_merge.rbegin();
		for (auto &left_it : left_ | std::ranges::views::reverse) {
			// check if there are any nodes left
			for (;split_merge_it->first >= left_it && split_merge_it != split_merge_begin; ++split_merge_it)
				// check if the node is split or merged
				if (split_merge_it->second == split_t) {
						--displacement; // decrement the displacement 
					} else
						++displacement; // decrement the displacement

			//check if we can exit
			if (split_merge_it == split_merge_begin && displacement == 0)
				break;

			left_it += displacement;
		}

		/* we than do the exact same thing with particule going to the right, rplacing ">=" at line 42 by a ">" */

		// finish first split 
		if (first_split) {
			rotate_once_left(left_);
			rotate_once_left(right_);
		}
	}
	//...
\end{lstlisting}

\subsubsection{hash()}

To hash graphs we simply combine all the positions of partricules going in one directions, a separator, the paricules going the other, a separator way and finally the hash of the graph's name.

\begin{lstlisting}[style=CStyle]
	//...
	size_t hash() const {
		if (hashed_)
			return hash_;

		hash_ = 0;
		boost::hash<unsigned int> hasher;

		// left hash 
		for (auto &l : left_)
			boost::hash_combine(hash_, l);

	 	 // separator 
	  	boost::hash_combine(hash_, separator);

	  	// right hash 
	  	for (auto &r : right_)
	   		boost::hash_combine(hash_, r);

	  	// separator 
	  	boost::hash_combine(hash_, separator);

	  	// name hash 
	  	boost::hash_combine(hash_, name_->hash());

	  hashed_ = true;
	  return hash_;
	}
	//...
}
\end{lstlisting}

\subsubsection{get\_split\_merges()}

To get the list of all indeces of splits and merges, we iterate through both the "left" and "right" positions, iterating the one that is the smallest so that it always converges towars positions that match (or are one apart for merges):

\begin{lstlisting}[style=CStyle]
/* function to get all split and merges */
std::vector<graph::split_merge_t> inline get_split_merge(graph_t* graph) {
	std::vector<graph::op_t> split_merge;

	unsigned int const last_idx = graph->size() - 1;

	auto left_it = graph->left().begin();
	auto right_it = graph->right().begin();
	bool first_or_last_split = *left_it == *right_it && *left_it == 0;
	first_or_last_split |= graph->left().back() == last_idx && graph->right().back() == last_idx;

	while (right_it < right_end = graph->right().end() && left_it < graph->left().end())
		if (*left_it == *right_it) {
			// check for split
			split_merge.push_back({*left_it, graph->split_t});

			++right_it;
			++left_it;
		} else if (*left_it < *right_it) {
			// chedk for merges
			if (*left_it == *right_it - 1)
				if (left_it == graph->left().end() - 1) {
					split_merge.push_back({*left_it, graph->merge_t});
				} else if (*(left_it + 1) != *right_it)
					split_merge.push_back({*left_it, graph->merge_t});

			++left_it;
		} else
			++right_it;

	if (!first_or_last_split)
		if (graph->left().back() == last_idx && graph->right()[0] == 0)
			split_merge.push_back({last_idx, graph->merge_t});

	return split_merge;
}
\end{lstlisting}

\subsection{superposition of graphs}

\subsubsection{class}

To allow parallel insertions of graphs into the lists of all graphs, we use tbb concurent hashmap. \\

We make the approximation that graphs with the same hash are equals and should therefor interact.

\begin{lstlisting}[style=CStyle]
class state {
	// hasher
	struct graph_hasher { /* */ }

	// comparator
	struct graph_comparator { /* */ }

	// type definition
	typedef tbb::concurrent_unordered_multimap<graph_t*, std::complex<long double>, graph_hasher, graph_comparator> graph_map_t;

	// main list 
	graph_map_t graphs_;

	// parameters
	std::complex<long double> non_merge_ = -1;
	std::complex<long double> merge_ = 0;

	//..
\end{lstlisting}

\subsubsection{reduce\_all()}

\begin{lstlisting}[style=CStyle]
	//...
	void reduce_all() {
		for(auto it = graphs_.begin(); it != graphs_.end();) {
			// range of similar graphs to delete
			auto upper = graphs_.equal_range(it->first).second;

			for(auto jt = std::next(it); jt != upper; ++jt)
		    	it->second += jt->second;

			// if the first graphgs has a zero probability, erase the whole range
			if (!check_zero(it->second))
				++it;

			it = graphs_.unsafe_erase(it, upper);
		}
	}
	//...
\end{lstlisting}

\subsubsection{step\_all()}

We pass a function "rule" that returns a vector of graphs and magnitude obtained from a single graph.

\begin{lstlisting}[style=CStyle]
	//...
	void state::step_all(std::function<tbb::concurrent_vector<std::pair<graph_t*, std::complex<long double>>>(graph_t* g)> rule) {
		graph_map_t buff;
		buff.swap(graphs_);

		#pragma omp parallel
		#pragma omp single
		for (auto & [graph, mag] : buff)
		#pragma omp task
		{
			auto graphs = rule(graph);
			for (auto & [graph_, mag_] : graphs)
			#pragma task
	  			graphs_.insert({graph_, mag_ * mag});
		}
	}
	//...
}
\end{lstlisting}

\subsubsection{source of errors}

\end{document}